---
title: "Homework-4"
author: "Madhu Peduri"
date: "6/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
    
## 1.A total of 30 Masters students in a Statistics course were asked to state how long they take to finish their weekly homework.The following is the reported time in hours: 2, 3, 1.5, 3.5, 4, 3, 1, 3, 2, 3.5, 4, 3, 2.5, 2.5, 2, 2, 1.5, 2, 2, 3, 3, 3, 1, 1, 1, 1.5, 2, 2.5, 2.5, 3

### (a). Write your own R code (function) to calculate the sample mean without using the mean() R function then confirm your answer using the mean() function.
\
We use below function to calculate the mean
\
\
$\bar{x} = \dfrac{1}{n}\sum\limits_{i=1}^{n}{x_{i}}$
\
```{r include = TRUE}
# Function to compute mean of a list of numbers
custoMean <- function(lst){
  n <- length(lst)
  nsum <- sum(lst)
  nmean <- nsum/n
  nmean
}
lst <- c(2, 3, 1.5, 3.5, 4, 3, 1, 3, 2, 3.5, 4, 3, 2.5, 2.5, 2, 2, 1.5, 2, 2, 3, 3, 3, 1, 1, 1, 1.5, 2, 2.5, 2.5, 3)
cmean <- custoMean(lst)
fmean <- mean(lst)
sprintf("Mean using our custom function is %.4f",cmean)
sprintf("Mean using the predefined function is %.4f",fmean)
```

### (b). Write your own R code (function) to calculate the sample variance without using the sd() and var() R functions. You can use the mean() and sum() functions, then confirm your answer using the var() function
\
We use below function to compute variance
\
\
$var = \dfrac{1}{n-1}(\sum\limits_{i=1}^{n}x_{i}^{2} - n\bar{x}^{2})$
\
```{r include = TRUE}
customVar <- function(lst){
  n <- length(lst)
  m <- custoMean(lst)
  res2 <- (lst - m)*(lst -m)
  nvar <- sum(res2)/(n-1)
  nvar
}
lst <- c(2, 3, 1.5, 3.5, 4, 3, 1, 3, 2, 3.5, 4, 3, 2.5, 2.5, 2, 2, 1.5, 2, 2, 3, 3, 3, 1, 1, 1, 1.5, 2, 2.5, 2.5, 3)
cvar <- customVar(lst)
fvar <- var(lst)
sprintf("Variance using our custom function is %.4f",cvar)
sprintf("Variance using the predefined function is %.4f",fvar)
```

## 2. Caleb wishes to take P = $10, 000 loan from a bank. The bank offers the loan at a monthly interest rate r = 3% for a period of n = 24 months. Calculate the monthly instalments m that Caleb will have to remit to the bank given that the principal is calculated as

We use below function to compute monthly interest
\
\
$P = m(\dfrac{1-(1+r)^{-n}}{r})$
\
$\Rightarrow m = P*\dfrac{r}{1 - (1+r)^{-n}}$
\
```{r include = TRUE}
compIntrest <- function(prin,rate,per){
  emi <- prin * (rate/(1-(1/((1+rate) ** per))))
  emi
}
mi = compIntrest(10000,0.03,24)
sprintf("Montly interest is %.4f",mi)
```