---
title: "HW6 - Exploratory Data Analysis (EDA)"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
#  html_document:
#    df_print: paged
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
library(ggplot2)
library(MASS)
library(Hmisc)
library(epiDisplay)
library(vcd)
library(mnormt)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

### 1.Use the built-in dataset cars.
#### 1.(a) Create a scatter plot of speed versus distance  

```{r}
car_df <- cars
plot(car_df$speed, 
     car_df$dist,
     type = "p", 
     col  = "blue",
     lwd  = 6, 
     xlab = "Speed (miles per hour)", 
     ylab = "Stopping Distance (feet)", 
     col.lab = "dark green",
     cex.lab = 1.5
     )
title(main = "Scatterplot of Speed versus Distance", 
      cex.main = 1.5,
      col.main = "red",
      sub = "Using plot() in R",
      cex.sub = 1
      )
axis(1,col.axis = "magenta")
axis(2,col.axis = "magenta")
```

#### 1.(b) Create Figure 2

```{r}
car_df <- cars
plot(car_df$speed, 
     car_df$dist,
     type = "p", 
     col  = "blue",
     lwd  = 6, 
     xlab = "speed", 
     ylab = "dist", 
     col.lab = "dark green",
     cex.lab = 1.5,
     axes = FALSE
     )
title(main = "Scatterplot of Speed versus Distance", 
      cex.main = 1.5,
      col.main = "red",
      sub = "Using plot() in R",
      cex.sub = 1
      )
legend("topleft",
       inset = 0.01,
       title = "Distance vs. Speed",
       c("Observation"),
       col=c("blue"), 
       pch=19,
       horiz=TRUE
       )
axis(1)
axis(2)
box(lty=3,col="red",lwd=2)
grid(lty=3)
abline(lm(dist ~ speed,data=car_df))
```

### 2. Use the economics built-in dataset and library ggplot2. Plot the time series of unemployment

```{r}
ggplot(economics, aes(date,unemploy)) +
       geom_line(size=0.8) +
       xlab('date') +
       ylab('Unemployed') +
       theme(panel.background = element_rect(fill = 'white', 
                                             colour = 'dark blue',
                                             size=3),
             panel.grid.major = element_line(colour = 'black', size = 0.2),
             panel.grid.minor = element_line(colour = 'light gray'),
             axis.ticks = element_line(size = 1),
             axis.text.x = element_text(size = 15,face="bold"),
             axis.text.y = element_text(size = 15,face="bold"),
             axis.title.x = element_text(size = 15,face="bold"),
             axis.title.y = element_text(size = 15,face="bold")
             )
```

### 3. Use the built-in dataset survey

#### 3.(a). Visualize survey dataset

```{r}
str(survey)
describe(survey)
des(survey)
summ(survey)
codebook(survey)
summary(survey)
```

### 3.(b) Generate frequency for features Sex, Exer and Smoke

```{r}
table(survey$Sex)
tab1(survey$Sex)
table(survey$Exer)
tab1(survey$Exer)
table(survey$Smoke)
tab1(survey$Smoke)
```

### 3.(c) Produce contingency tables

```{r}
table(survey$Sex, survey$Exer, useNA = "ifany")
chisq.test(survey$Sex, survey$Exer, correct=FALSE)

table(survey$Smoke, survey$Exer, useNA = "ifany")
chisq.test(survey$Smoke, survey$Exer, correct=FALSE)

table(survey$Smoke, survey$Sex, useNA = "ifany")
chisq.test(survey$Smoke, survey$Sex, correct=FALSE)
```

We can make below observation,

- We have p-value = 0.5 from chi-test between features Somke and Exercies. This is high compared to significant value 0.05 and we can say, there is high correlation between these features.
- We have p-value = 0.3 from chi-test between features Somke and Sex. This is high compared to significant value 0.05 and we can say, there is correlation between these features.
- We have p-value = 0.06 from chi-test between features somke and Sex. This is not high  and we can say, there is correlation between these features 

### 3.(d) Correlation matrix

```{r}
data("survey")
ff <- lm(Height ~ Wr.Hnd, data = survey)
summary(ff)
# calculation of Pearson' correlation coefficient
cor(survey$Wr.Hnd, survey$Height, use = "complete")

# This code was used to produce the correlation
# matrix
library(psych)
dat0 <- survey[, c("Pulse", "Age", "Height", "NW.Hnd",
"Wr.Hnd")]
pairs.panels(dat0)

summary(lm(NW.Hnd ~ Wr.Hnd, data = survey))
```

We can make below observations,

- Between features, hand span and hight, we have correlation factor as 0.6. By this we can say, there is medium level of correlation, 1 being perfecly correlated.
- We have R-squared = 0.361. From this we can say that the linear regression predictions of the feature height using hand-span are not good. Low R-square suggest that, there errors are hight between actual and predictions. This is substantiated by the fact that correlation between these two is not high. 
- From the Correlation matrix we can say, feature pair 'Wr.Hnd' and 'NW.Hnd' are hightly correlated with 0.95 value. R-squared for the linear regression between these two features is 0.95 which substantiates the correlation. 

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
