---
title: "HW8 - Simple Regression"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
library(ggplot2)
library(MASS)
library(Hmisc)
library(epiDisplay)
library(vcd)
library(mnormt)
library(MASS)
library(car)
library(ggpubr)
library(PairedData)
library(lmtest)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

### 1. Goodness of Fit

#### 1.1 Generate a vector x containing 100 random numbers from a standard normal distribution and visualize the data:

```{r}
x <- rnorm(100)
par(mfrow = c(2, 2))
plot(x, las = 1, cex = 1.5)
hist(x, las = 1, cex = 1.5)
boxplot(x, las = 1, cex = 1.5)
```

#### 1.2 To induce a linear relationship between x and a dependent variable y, generate the vector y using a linear transformation of the vector x:

```{r}
y1 <- 0 + 1 * x ## A perfect linear association
y2 <- 0 + 1 * x + rnorm(length(x), mean = 0, sd = sqrt(1)) # Add a little N(0,1) noise (error)
```

#### 1.3 Compute, y3,y4,y5,y6
#### $y_{i} = \beta_{0} + \beta_{1}x_{i} + \epsilon_{i} ~ N(0,\sigma^{2})$

```{r}
y3 <- 0 + 1 * x + rnorm(length(x), mean = 0, sd = sqrt(3))
y4 <- 0 + 1 * x + rnorm(length(x), mean = 0, sd = sqrt(5))
y5 <- 0 + 1 * x + rnorm(length(x), mean = 0, sd = sqrt(7))
y6 <- 0 + 1 * x + rnorm(length(x), mean = 0, sd = sqrt(9))
```

#### 1.4 Generate scatterplots, r, and R2

```{r}
vlist <- list(y1,y2,y3,y4,y5,y6)
par(mfrow = c(2, 3))
nv <- c(1,2,3,4,5,6)
rv <- c(1,2,3,4,5,6)
sdv <- c(sqrt(0),sqrt(1),sqrt(3),sqrt(5),sqrt(7),sqrt(9))
R2v <- c(1,2,3,4,5,6)
i=0
for (yi in vlist){
  i <- i + 1
  rv[i] <- cor(x,yi)
  R2v[i] <- summary(lm(yi ~ x))$r.squared
  plot(x, yi, las = 1, cex = 1.5,ylab=paste('y',nv[i],sep=''))
  mtext(bquote(paste(R^2 == .(R2v[i]), ", ", ~r == .(rv[i]))), col = "blue")
}
```

#### 1.5 Plot r and R2 versus standard deviation

```{r}
par(mfrow = c(1, 2))
plot(sdv,rv,type="l",xlab="std dv", ylab="Correlation")
plot(sdv,R2v,type = "l",xlab="std dv", ylab="R Squared")
```


#### Observation
- From above plots, we can say that as standard deviation of the noise increases, the coefficients of correlation and R-squared decreases. 

### 2. Simple Linear Regression

#### 1. Write out the regression model that would explore the proposed relationship and state the model assumptions.

```{r}
gscore <- read.csv("GPA.csv")
regmodel <- lm(gscore$GPA ~ gscore$ACT)
modelsumm <- summary(regmodel)
names(regmodel)
names(modelsumm)
modelsumm
```

#### Assumptions
- Homoscedasticity
- Normality
- Correlation

#### 2. Testing Assumptions

```{r}
# Homoscedasticity
par(mfrow = c(1, 2))
plot(fitted(regmodel), residuals(regmodel))
abline(h = 0)
plot(fitted(regmodel), abs(residuals(regmodel)))
summary(lm(abs(residuals(regmodel)) ~ fitted(regmodel)))
ncvTest(regmodel)

# Normality
qqnorm(residuals(regmodel), main = "QQ Plot")
qqline(residuals(regmodel))
#ggqqplot(residuals(regmodel))
shapiro.test(residuals(regmodel))

# Correlation
plot(residuals(regmodel), main="correlation")
abline(h=0)
#summary(lm(residuals(regmodel)[-1] ~ -1+residuals(regmodel)[-541]))
dwtest(regmodel)
```

#### Observations
Homoscedasticity - By observing the plot between fitted and residuals, they look like non-constant variance plots. P-value = 0.4 (0.05) from ncvTest suggest non-constant variance.

Normality - By observing the plots of qqnorm, we can say that all points are along the reference lines and that suggests the normality. However, Shapiro-wilk test gives a p-value < 0.05 significant value. This shows non-normality nature of the distribution of residuals. Normality is the least worrisome assumption.

Correlation - By observing the correlation plot, we can say th residuals are uncorrelated. dwtest has p-value 0.2 (>0.05) greater than significant value also suggest uncorrelation

#### 3. Intercepts

Intercept: $\beta_{0} = 2.1140$ \
Y-intercept: $\beta_{1} = 0.0388$\
Regression function: $GPA = 2.1140 + 0.0388*ACT$

#### 4. Plot raw data

```{r}
gscore$predict <- lapply(gscore$GPA,function(x){2.1140 + 0.0388*x})
plot(gscore$GPA, gscore$ACT,xlab='GPA',ylab='ACT',yaxt = "n")
par(new = TRUE)
plot(gscore$GPA, gscore$predict, type='l', col='blue',xlab='',ylab='')
abline(regmodel,col='red',lty=3)
title("raw data with estimate function")
```

#### 5. Percentage of variation

For every 1 mark increase in ACT score, the GPA function increases by 0.04%.

#### 6. Confidence interval

```{r}
confint(regmodel)
```

#### Observations
- 95% confidence interval for $\beta_{1}=$ 0.04
- (95% CI: 0.04, 0.01)
- From the confidence interval (2.5 to 97.5) above, we can say zero is not included in it. This says that there is evidence of a linear relationship between predictor GPA and response ACT in the sample. 

#### 7. Test linear association 

- Null hypothesis of linear regression $H_{0}: \beta{1} = 0$ that says, that predictor has not effect on the output. 
- Alternate hypothesis $H_{1}:\beta_{1} \neq 0$ that says, there is a linear relation between predictor and the output.
- We can see from the confidence interval do not include zero. That implies at the evidence of linear relation ship between Predictor GPA and output ACT score.

#### 8. p-value of linear regression

- We can see from the summary of regression model, that p-value = 0.00292. This is less than the significant value 0.01. Using this we can say that alternate hypothesis is true and shows evidence of linear relationship between GPA and ACT score.

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
