---
title: "HW12 - Poisson Regression"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
  html_document:
    df_print: paged
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
library(ggplot2)
library(MASS)
library(Hmisc)
library(epiDisplay)
library(vcd)
library(mnormt)
library(MASS)
library(car)
library(ggpubr)
library(PairedData)
library(lmtest)
library(xtable)
library(faraway)
library(leaps)
library(psych)
library(Matrix)
library(dobson)
library(jtools)
library(Rcpp)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

### Problem 1  

```{r}
# Convert the doctors dataframe to appropriate format
doc <- data.frame(
  age       = as.numeric(unclass(as.factor(doctors$age))), 
  agesq     = (as.numeric(unclass(as.factor(doctors$age))))^2,
  agecat    = as.character(doctors$age),
  smoke     = as.numeric(unclass(as.factor(doctors$smoking))),
  deaths    = as.numeric(doctors$deaths),
  personyrs = as.numeric(doctors$'person-years')
)
doc
des(doc)
```

### Problem 1(a)

```{r}
# GLM model 
poismod <- glm(deaths ~ age + agesq + smoke + smoke:age, offset = log(personyrs), family = poisson(link = "log"), data = doc)

sumpois <- summary(poismod)
sumpois
summpois <- summ(poismod, confint = TRUE, digits = 3, ci.width = 0.95)

sprintf("Residual deviance of the poisson model : %.3f",sumpois$deviance)
sprintf("Residual degrees of freedom of the poisson model : %.3f",sumpois$df.residual)
print("Pseudo R2 (Mcfadden) of the model : 0.943")
```

### Problem 1(b)

```{r}
sprintf("Null deviance of the model : %.3f",poismod$null.deviance)
sprintf("Deviance of the model residuals : %.3f",poismod$deviance)

devresd <- round(residuals(poismod, type = "deviance"), 3)
devresd

stddevres <- rstandard(poismod)
stddevres

sprintf("Chi-square goodness-of-fit value : %.3f",sum(devresd^2))

par(mfrow = c(1, 2))
plot(log(predict(poismod)),devresd, ylab="residuals")
abline(h=0)
plot(log(predict(poismod)),stddevres, ylab="Standardized residuals")
abline(h=0)

# standardized Pearson's residuals
pearson.resid <- round(resid(poismod, type = "pearson"), 3)
pearson.resid

stdpearsons <- rstandard(poismod, type = "pearson")
stdpearsons

sprintf("Pearson goodness-of-fit value : %.3f",sum(pearson.resid^2))

par(mfrow = c(1, 2))
plot(log(predict(poismod)),pearson.resid, ylab="pearson residuals")
abline(h=0)
plot(log(predict(poismod)),stdpearsons, ylab="pearson Standardized residuals")
abline(h=0)

# expected <- fitted.values(docm1, type='response')
expected <- round(predict(poismod, type = "response"), 2)
expected

par(mfrow = c(1, 2))
plot(doc$deaths, ylab="Actual")
plot(expected, ylab="Predicted")
```

#### Observations
- Outliers will be 2 standard deviations away from mean. We do not have standard residuals more than +2 or -2 suggesting no outliers. 
- We do not see much difference between actual and predicted values. 

### Problem 1(c)

```{r}
kable(cbind(doc[, c(1, 3, 4, 5)], expected, pearson.resid, devresd))
```

### Problem 1(d)

```{r}
# Chi square
pchisq(deviance(poismod), df.residual(poismod), lower = F)
deviance(poismod)
pr <- residuals(poismod,"pearson")
sum(pr^2)

# poisson goodness of fit values
poisgof(poismod)

# plots
par(mfrow = c(2, 2))
plot(log(fitted(poismod)),(doc$deaths-fitted(poismod))^2,xlab = expression(hat(mu)),ylab = expression((y-hat(mu))^2))
abline(0,1)
halfnorm(residuals(poismod))
shapiro.qqnorm(residuals(poismod, type = "deviance"))
influencePlot(poismod)

# Confidence intervals of each predictor
kable(cbind(summary(poismod)$coef, confint(poismod)))

# Significance of each predictor relative to full model
kable(drop1(poismod, test = "F"))

# Goodness of fit measures using anova
kable(anova(poismod, test = "Chisq"))

# mean and variance
phi <- sum(pr^2)/df.residual(poismod)
round(c(phi,sqrt(phi)),4)
mean(fitted(poismod))
sqrt(var(fitted(poismod)))
```

#### Observations
- Deviance: This explains the un-explained variance in our data. Lesser the deviance more good the model is. We have a less deviance of 1.64 suggesting good explainability of the variance.
- Chisquare: It gives an idea of how much the fitted values differ from the expected values. Lesser the chisquare more dood the model is. We have a less chisquare of 0.897 suggesting good fitted values.
- poisgof: The null hypothesis of this is a good goodness-of-fit measure for the poisson regression model. We have the p-value 0.897 > 0.05 significant value suggesting good fit of the model. 
- Plots: We can see the normality of the residuals. 
- drop1: We can see less AIC value for the predictor age*smoke suggesting it as good predictor.
- Anova: We can see the Pr values are small for all the predictors. 
- Assumption of the poission model is that mean and variance of the fitted values should be same. We can see they are approximately near suggesting a good model. 

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
