---
title: "Homework-2"
author: "Madhu Peduri"
date: "6/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
    
## 1.(a) Create a vector of length 10 and fill it with a sequence of integers 
  
We use concatenat function to create a vector object, 
\
```{r include = TRUE}
v <- c(1:10)
length(v)
```
\    
$\Rightarrow$ Vector = `r v` 

## 1.(b) Coerce the vector into a matrix of 5 rows and 2 columns.

We use matrix and as.matrix() functions to coerce a vector in to matrix.
\
```{r include = TRUE}
m <- as.matrix(matrix(v,nrow=5,ncol=2))
print(m)
print(class(m))
```

## 1.(c) Name the columns of the matrix "A" and "B".

We use dimnames() function to add names to the columns of above matrix
\
```{r include = TRUE}
dimnames(m)[2] <- list(c('A','B'))
print(m)
```

## 2.Assume that we have recorded the names and ages for four people: James is 27, Art is 42, Kate is 29, and Alex is 33.

### (a) Create a vector of names and a vector of ages from the data, making sure that you keep the ordering of the elements consistent.

\
```{r include = TRUE}
nameVec <- c('James', 'Art', 'Kate', 'Alex')
ageVec <- c(27,42,29,33)
```
\
$\Rightarrow$ Names = `r nameVec` \
$\Rightarrow$ Ages = `r ageVec` 

### (b) Use the class() function to print the class metadata R has stored for each vector.

\
```{r include = TRUE}
cn <- class(nameVec)
ca <- class(ageVec)
```
\
$\Rightarrow$ Class of Names Vector = `r cn` \
$\Rightarrow$ Class of Age vector = `r ca` 

### (c) Using data.frame(), combine the two vectors into a dataframe and name the columns something informative. Print out the dataframe.

\
```{r include = TRUE}
df <- data.frame(names=nameVec, ages = ageVec)
print(df)
```

## 3.A hypothetical data with heights (in cm) and weights (in kg) of 10 family members are given

### (a) Create a vector called ‘ht’ corresponding to the heights of the 10 family members. Assign the names of the family members to the ‘names’ attribute of this vector.

\
```{r include = TRUE}
ht <- c(120,172,163,158,153,148,160,170,155,167)
names(ht) <- c('Niece','Son','GrandPa','Daughter','Yai','GrandMa','Aunty','Uncle','Mom','Dad')
print(ht)
```
### (b) Create a vector called ‘wt’ corresponding to the family member’s weights.

\
```{r include = TRUE}
wt <- c(22,52,71,51,51,60,50,67,53,64)
print(wt)
```
### (c) Compute the body mass index (BMI: units should be kg/m2) of each person where BMI = weight/height2.

\
```{r include = TRUE}
bmi <- mapply(function(x,y){(y * 100 * 100)/(x * x)}, ht,wt)
print(bmi) 
```
### (d) Identify the persons who have the lowest and highest BMI and calculate the standard deviation of the BMI.

\
```{r include = TRUE}
minBmi <- names(bmi)[bmi == min(bmi)]
maxBmi <- names(bmi)[bmi == max(bmi)]
sdBmi  <- sd(bmi)
```
\
$\Rightarrow$ Minimum BMI = `r minBmi` \
$\Rightarrow$ Maximum BMI = `r maxBmi` \
$\Rightarrow$ Standard Deviation in BMI = `r sdBmi`
