---
title: "Homework-3"
author: "Madhu Peduri"
date: "6/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
    
## 1.(a) Clean up the workspace using the rm() function. Use the data() function to display the built-in datasets you can access. Use the R help to learn more about the ‘longley’ dataset: ?longley. 

We use rm() function to remove objects to control the usage of memory.
We use data() to load the 'longley' dataset and head() to see the sample data.
\
```{r include = TRUE}
rm(list = ls(all=TRUE))
ldata = longley
head(ldata)
class(ldata)
names(ldata)
```

## 1.(b) Print only the records in the ‘longley’ dataset that are from the years 1947-1950

```{r include = TRUE}
longley[longley$Year == 1947:1950,]
```

## 1.(c) Plot (Unemployed ~ Year)

```{r include = TRUE}
xmin = min(longley$Year)
xmax = max(longley$Year)
plot(longley$Year,longley$Unemployed,xlab = 'Year',ylab='Unemployed',xaxt="n",xlim=c(xmin,xmax))
axis(1, at = xmin:xmax)
```

## 1.(d) Change the type of plot to a line

```{r include = TRUE}
xmin = min(longley$Year)
xmax = max(longley$Year)
plot(longley$Year,longley$Unemployed,type='l',xlab = 'Year',ylab='Unemployed',xaxt="n",xlim=c(xmin,xmax))
axis(1, at = xmin:xmax)
```

## 2. You track your commute times for two weeks and record the following (in minutes):17 16 20 24 22 15 21 15 17 22.

### a. Enter these numbers into R and find the 5-number summary.

```{r include = TRUE}
ctime <- c(17, 16, 20, 24, 22, 15, 21, 15, 17, 22)
summary(ctime)
```

### b. You find a data entry error, the number 24 should have been 18. Using R, replace the incorrect value without reentering the entire set of data and find the new 5-number summary.

```{r include = TRUE}
ctime <- c(17, 16, 20, 24, 22, 15, 21, 15, 17, 22)
ctime[ctime == 24] <- 18
print(ctime)
summary(ctime)
```

### c. Use R to count the number of times your commute was at least 20 minutes.

```{r include = TRUE}
lt <- length(ctime[ctime <= 20])
sprintf("Number of times commute was atleast 20 min: %i", lt)
```

### d. Use R to calculate the percent of your commutes that were less than 17 minutes.

```{r include = TRUE}
lt <- length(ctime[ctime < 17])
sprintf("Number of times commute was less than 17 min: %i", lt)
```

## 3. Using the maltreat.dta dataset, explore the variable ethnic using tab1(ethnic). There are spelling mistakes that need to be corrected. Correct mis-spelt names, and create a numeric, categorical variable ethncity. The “Jola” cleaning code for part (i) has been provided. Finish the remaining part of the code and produce the final (clean) bar chart.

```{r include = TRUE}
library("readstata13")
library("epiDisplay")

#Read the maltread data
maltreat <- read.dta13("data/maltreat.dta")

#Display the frequency
tab1(maltreat$ethnic, col = "grey")

#create the categorical feature enthnicity from enthnic
maltreat$ethnicity <- as.factor(maltreat$ethnic)
levels(maltreat$ethnicity)
```

### (i). Replace ethnic = “Jola” if ethnic value starts with a “J”.

```{r include = TRUE}
#Correct Jola
levels(maltreat$ethnicity)[startsWith(levels(maltreat$ethnicity),
"J")] <- "Jola"
maltreat$ethnicity[startsWith(as.character(maltreat$ethnicity),"J")] <- "Jola"
```

### (ii). Replace ethnic = “Mandinka” if ethnic value starts with an “M”

```{r include = TRUE}
#Correct Jola
levels(maltreat$ethnicity)[startsWith(levels(maltreat$ethnicity),
"M")] <- "Mandinka"
maltreat$ethnicity[startsWith(as.character(maltreat$ethnicity),"M")] <- "Mandinka"
```

### (iii). Replace ethnic = “Serahule” if ethnic value starts with an “S”

```{r include = TRUE}
#Correct Jola
levels(maltreat$ethnicity)[startsWith(levels(maltreat$ethnicity),
"S")] <- "Serahule"
maltreat$ethnicity[startsWith(as.character(maltreat$ethnicity),"S")] <- "Serahule"
```

### (iv). Replace ethnic = “Wollof” if ethnic value starts with a “W”

```{r include = TRUE}
#Correct Jola
levels(maltreat$ethnicity)[startsWith(levels(maltreat$ethnicity),
"W")] <- "Wollof"
maltreat$ethnicity[startsWith(as.character(maltreat$ethnicity),"W")] <- "Wollof"

# Correct the feature enthic from corrected enthnicity
maltreat$ethnic <- maltreat$ethnicity
levels(maltreat$ethnic) <- levels(maltreat$ethnicity)
tab1(maltreat$ethnic, col = "grey")

# Change the feature enthnicity to numeric and categorical
maltreat$ethnicity <- as.numeric(factor(maltreat$ethnicity,levels=unique(maltreat$ethnic),exclude = NULL))
tab1(maltreat$ethnicity,col = "grey")
```