---
title: "HW10 - Matrix Approach to Linear Regression"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
  html_document:
    df_print: paged
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
library(ggplot2)
library(MASS)
library(Hmisc)
library(epiDisplay)
library(vcd)
library(mnormt)
library(MASS)
library(car)
library(ggpubr)
library(PairedData)
library(lmtest)
library(xtable)
library(faraway)
library(leaps)
library(psych)
library(Matrix)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

### Problem 1(a) 

```{r}
A = matrix(c(1,5,0,1,0,5,1,0,5), byrow=TRUE, ncol=3)

rankA <- rankMatrix(A)
sprintf("Rank of matrix A is %i",rankA)

detA <- det(A)
sprintf("Determinant of matrix A is %i",detA)

traceA <- sum(diag(A))
sprintf("Trace of matrix A is %i",traceA)

A
```

(i) are the column vectors of A linearly dependent?
- Matrix A has row 2 and row 3 same, which are linearly dependent. We have determinant = 0, which makes it singular suggesting linear dependency.

(ii) What is the rank of A?
- Rank of a matrix A is 2

(iii) What must be the determinant of A?
- Determinant of Matrix A is zero

(iv) Find the trace of A.
- Trace of matrix A is 6

### Problem 1(b)

```{r}
B = matrix(c(5,1,3,4,0,5,1,9,6), byrow=TRUE, ncol=3)

rankB <- rankMatrix(B)
sprintf("Rank of matrix B is %i",rankB)

detB <- det(B)
sprintf("Determinant of matrix B is %.2f",detB)

traceB <- sum(diag(B))
sprintf("Trace of matrix B is %i",traceB)

inverseB <- round(solve(B),3)
inverseB
```

(i) are the column vectors of B linearly dependent?
- Matrix B do not have any rows or columns that can be derviced from one another. Determinant of matrix is not zero that makes it non-singular and of full-rank suggesting no linear dependency. 
(ii) What is the rank of B?
- Rank of the matrix B is 3 which is equal to its largest dimension. 
(iii) Calculate the determinant of B.
- Determinant of matrix B = -136.
(iv) Find the trace of B.
- Trace of matrix B = 11
(v) Find the inverse of B,
$B^{-1} = \begin{bmatrix} 0.331 & -0.154 & -0.037 \\ 0.140 & -0.199 &  0.096 \\ -0.265 &  0.324 &  0.029\end{bmatrix}$

### Problem 2

```{r}

c1 <- c(16,5,10,15,13,22)
c2 <- c(4,1,2,3,3,4)
data <- cbind(c1,c2)

Y <- matrix(c1,byrow = TRUE,ncol=1)
X <- matrix(cbind(1,c2), ncol = 2)

# (i)
yty <- t(Y)%*%Y
yty

# (ii)
xtx <- t(X)%*%X
xtx

# (iii)
xty <- t(X)%*%Y
xty

# (iv)
xtxInv <- solve(xtx)
xtxInv

# (v)
coef_beta <- xtxInv%*%xty
coef_beta

# (vi)
Yhat <- X%*%coef_beta
Yhat
e_residual <- (Y - Yhat)
res_vec <- as.vector(e_residual)
res_vec

# (vii)
er <- (Yhat - mean(Y))^2
SSR <- sum(er)
sprintf("Sum of squares due to regression is:%.3f",SSR)

# (viii)
er <- (Y - Yhat)^2
SSE <- sum(er)
sprintf("Sum of squares of residuals is:%.3f",SSE)

# (ix) Variance Covariance Matrix of coefficients 
MSE <- SSE/(length(Y) - 2)
varCovarMatrix <- MSE * xtxInv
varCovarMatrix
```

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
