---
title: "HW14 - Debug"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
  html_document:
    df_print: paged
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
library(ggplot2)
library(MASS)
library(Hmisc)
library(epiDisplay)
library(vcd)
library(mnormt)
library(MASS)
library(car)
library(ggpubr)
library(PairedData)
library(lmtest)
library(xtable)
library(faraway)
library(leaps)
library(psych)
library(Matrix)
library(dobson)
library(jtools)
library(Rcpp)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

### Q1(a)

```{r}
calculate.exp <- function(my.number){
exp.num <- (-my.number) ^ 2
print(paste("Value of exp.num :",exp.num))
result <- exp(exp.num)
return(result)
}
```

### Q1(b)

```{r}
calculate.exp(1)
```

### Q1(c)

```{r}
debugonce(calculate.exp)
calculate.exp(1)
```

### Q1(d)
- Debug: This method can be used to debug a function by setting a debugging flag. When a function flagged for debugging is entered, normal execution is suspended and the body of function is executed one statement at a time. A new browser context is initiated for each step. Once we flag a function using this, this remains for the entire session, that is, every call to this function launches the debug browser context.
- Debugonce: This method is same as debug(), but it launches only for immediate next function call and further calls will be run in normal execution mode.

### Q1(e)

```{r}
calculate.exp <- function(my.number){
exp.num <- -(my.number) ^ 2
browser()
print(paste("Value of exp.num :",exp.num))
result <- exp(exp.num)
return(result)
}
calculate.exp(1)
```

### Q1(f)
- Editor breakpoints: Another common way to stop on a line of code is to set a breakpoint on that line. We can do this in RStudio by clicking to the left of the line number in the editor, or by pressing Shift+F9 with your cursor on the desired line.
- We call this an “editor breakpoint”. Editor breakpoints take effect immediately and don’t require you to change your code (unlike browser() breakpoints, below).

### Q2

```{r}
my.simulation <- function(mu){
# initiate an empty value
mean.vec <- NA
for (i in 1:100000){
# Step 1
simu.data <- rnorm(n = 10, mean = mu, sd = 1)
# Step 2
mean.simu <- mean(simu.data)
# Step 3
mean.vec <- c(mean.vec, mean.simu)
}
result <- mean(mean.vec)
return(result)
}
# print the time it takes to execute the function
system.time(print(my.simulation(10)))
```

### Q2 (a)

```{r}
my.simulation <- function(mu){
# initiate an empty value
mean.vec <- NA
for (i in 1:10000){
# Step 1
simu.data <- rnorm(n = 10, mean = mu, sd = 1)
# Step 2
mean.simu <- mean(simu.data)
# Step 3
mean.vec <- c(mean.vec, mean.simu)
}
result <- mean(mean.vec)
return(result)
}
# print the time it takes to execute the function
system.time(print(my.simulation(10)))
```

### Q2 (b)

```{r}
my.simulation <- function(mu){
# initiate an empty value
mean.vec <- 0 ## Corrected statement
for (i in 1:100){
# Step 1
simu.data <- rnorm(n = 10, mean = mu, sd = 1)
# Step 2
mean.simu <- mean(simu.data)
# Step 3
mean.vec <- c(mean.vec, mean.simu)
}
result <- mean(mean.vec)
return(result)
}
# print the time it takes to execute the function
system.time(print(my.simulation(10)))
```

### Q3 (a) 

```{r}
get.runs <- function(x, k){
n <- length(x)
runs <- NULL
for (i in 1:(n-k)){
  if(all(x[i:(i+k-1)] == TRUE)){
    runs <- c(runs, i)
    }
  }
return(runs)
}
```

### Q3 (b)

```{r}
# this function below should return a vector
# of (4,5,8) because there are a run of two TRUEs
# in those indices.
get.runs(c(TRUE, FALSE, FALSE, TRUE, TRUE, TRUE, FALSE, TRUE, TRUE), 2)
```

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
