---
title: "HW7 - Statistical Inference"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
#  html_document:
#    df_print: paged
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
library(ggplot2)
library(MASS)
library(Hmisc)
library(epiDisplay)
library(vcd)
library(mnormt)
library(MASS)
library(car)
library(ggpubr)
library(PairedData)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

### 1.Analyzing Data with a Categorical Outcome.
#### 1.1 (a) Enter the data into R in an expanded form

```{r}
cat <- c(rep('Yes',630), rep('No',670))
vec <- factor(cat)
binge <- data.frame(vec)
names(binge) <- c('BingeDrink')
head(binge)
```

#### 1.1 (b) Frequency charts

```{r}
par(mfrow=c(1,2))
t1 <- tab1(binge,main='')
t2 <- tab1(binge,bar.values = 'percent',main='')
mtext('Binge Drinking on Campus:',side=3,line=2.4,at=-0.8,cex=1.4)
mtext('A Survey of N = 1300 Undergraduates',line=1,side=3,at=-0.8,cex=1.4)
```

#### 1.2 (a) Prop & Binomial test

```{r}
prop.test(x=length(binge$BingeDrink[binge$BingeDrink == 'Yes']),n=length(binge$BingeDrink),p=0.5)
binom.test(x=length(binge$BingeDrink[binge$BingeDrink == 'Yes']),n=length(binge$BingeDrink),p=0.5)
```
#### Observations:
- A binomial test compares a sample proportion to a hypothesized proportion. The test has the following null and alternative hypothesis. 
- We have 630 number of successes in an experiment with total number of 1300 trials with a probability of 1/2 for a success for a given trial.
- We obtained a p-value = 0.3 (> 0.05) in our binomial test, which suggests that we can accept the null hypothesis and conclude that we got 'Yes' in our experiment. 

#### 1.2 (b)(i) Chi-square goodness

```{r}
# When probability of happeining Yes and No are equal to 1/2
prob <- c(length(binge$BingeDrink[binge$BingeDrink == 'Yes']),length(binge$BingeDrink[binge$BingeDrink == 'No']))
chisq.test(prob, p = c(1/2, 1/2))
chisq.test(prob, p = c(1/3, 2/3))
```

#### Observations:
- The chi-square goodness of fit test is used to compare the observed distribution to an expected distribution, in a situation where we have two or more categories in a discrete data. In other words, it compares multiple observed proportions to expected probabilities.
- For the hypothesis $H_{0}:p_{1} = p_{2}$, with same probability for both categorical values, we have p-value = 0.3 (> 0.05). This says that, observed proportions are not different from the expected proportions.
- For the hypothesis $H_{1}:p_{1} \neq p_{2}$, with different probability for both categorical values, we have p-value very less. This says that, observed proportions are different from the expected proportions.

#### 1.2 (b)(ii)

```{r}
pi = 0.247
chisq.test(prob, p = c(pi, 1-pi))
```

#### Observations:
- If we use the proportion = 0.247 by NSDUH as expected hypothesis, we are getting a less p-value. We can say this, our actual proportion do not comply with expected hypothesis.
- We can see 0.247*1300 = 321.1 (< 630) which is less than the actual number of successes. 

#### 1.2 (c)

```{r}
prop.test(x = 630, n = 1300, p = 0.247, alternative = "g")
binom.test(x = 630, n = 1300, p = 0.247, alternative = "g")
```

#### Observations:
- We can see in both the cases we have p-value is less. 
- If we have proportion = 0.5 (>0.247), we have p-value that is sufficient which can be equivalent to hypothesis.

#### 1.3(a) Read Sales Data

```{r}
sales <- read.csv("sales.csv")
tabs <- xtabs(~Region + Sport, data = sales)
tabs
```

#### 1.3(b) Mosaic Plot

```{r}
ptab <- prop.table(tabs)
ptab

mosaic(ptab,zero_size = 0)
```

#### Observations:
- We can observe that Mosaic plot shows the equivalent densities. 
- We can see that, Region B has highes sales for the Summer sport gear.

#### 1.3(c) Chi-square test of independence

```{r}
chisq.test(tabs)
```

#### Observations:
- We can see very less p-value which shows the less dependency between Sales of sports gear and geographic region.

### 2. Analysis of Continuous Outcome Data

#### 2.1(a) One-Sample Tests

```{r}
car <- c(19, 26, 24, 21, 24, 23, 26, 24, 23, 20, 21, 24, 18, 21, 20, 23, 24, 26, 25, 19, 24, 23, 27, 24, 26, 25, 20, 21, 19, 23)
ttest <- t.test(car, mu=25)
sprintf("p-value of t-test: %.10f",ttest$p.value)
sprintf("Mean of the sample: %.2f", mean(car))
sprintf("Standard deviation of the sample: %.2f", sd(car))

plot(car, type="l", xlab="trial", ylab="mpg", main="mileage")
par(new = TRUE)
abline(h=mean(car), col="blue",lty = 3)
```

#### Observations:
- We can see p-value of our t-test is less (inferior) to the significant value 0.05. This suggest that null hypothesis is incorrect and the hypothesis $H_{1}: \mu \neq 25$ is correct. 

#### 2.1(b) Shaprio test

```{r}
qqnorm(car, pch = 1, frame = FALSE)
qqline(car, col = "steelblue", lwd = 2)
ggqqplot(car)
shapirotest <- shapiro.test(car)
sprintf("P-value of Shapiro normality test:%.4f",shapirotest$p.value)
```

#### Observations:
- From qqplots, we can see the sample quantiles follow a straight line and fall with in the range. 
- The p-value of shapiro.test is 0.09 which is greater than significant value 0.05 suggests that our data follow normal distribution.

#### 2.1(c) t - test

```{r}
ttest <- t.test(car, alternative='less', mu=25)
ttest
sprintf("P-value of t-test test:%.5f",ttest$p.value)
```

#### Observations:
- There is insufficient evidence to conclude that the average fuel economy of 2018 Sedans is less than the 25 mpg reported by Company A (p = 0.00002, 95%CI : -Inf 23.5).

### 2.2(a) Dependent Samples Tests

```{r}
id <- c(1:10)
pre <- c(899.63, 913.51, 897.05, 889.18, 903.2, 916.06, 899.08, 892.75, 901.47, 902.63)
post <- c(899.53, 899.38, 879.25, 867.35, 897.97, 921.42, 895.52, 893.95, 889.44, 898.14)
datap <- data.frame(cbind(id, pre, post))
paired.plotProfiles(datap, "pre", "post") + 
  geom_line(color="blue") +
  theme(panel.background = element_rect(fill = 'white', 
                                             colour = 'dark blue',
                                             size=3),
        axis.text.x = element_text(size = 15,face="bold"),
        axis.text.y = element_text(size = 15,face="bold"),
        panel.grid.major = element_line(colour = 'light gray', size = 0.4),
        panel.grid.minor = element_line(colour = 'light gray', size = 0.4)
       )
```

### 2.2(b) Dependent Samples Tests
#### (i) Box plots

```{r}
sprintf("Pre-swim time - Mean:%.2f and Standard deviation:%.2f",mean(datap$pre),sd(datap$pre))
sprintf("Post-swim time - Mean:%.2f and Standard deviation:%.2f",mean(datap$post),sd(datap$post))
par(mfrow=c(1,2))
boxplot(x=datap$pre,xlab='pre',ylab='times')
boxplot(x=datap$post,xlab='post')
```

#### (ii) Box plots

```{r}
paired_dif <- datap$pre - datap$post
sprintf("Paired differences - Mean:%.2f and Standard deviation:%.2f",mean(paired_dif),sd(paired_dif))
boxplot(x=paired_dif,xlab='paired diff',ylab='times')
```

#### (iii) Normality test

```{r}
qqnorm(paired_dif, pch = 1, frame = FALSE)
qqline(paired_dif, col = "steelblue", lwd = 2)
ggqqplot(paired_dif)
shapirotest <- shapiro.test(paired_dif)
sprintf("P-value of Shapiro normality test:%.4f",shapirotest$p.value)
```

#### Observation
- From the plots and shapiro test's p-value = 0.78 indicates the nature of normality in paired differences

#### (iii) t - test

```{r}
datap$diff <- pre - post
t.test(pre, post, paired = TRUE, alternative = "greater")
t.test(datap$diff, alternative = "greater",mu=0)
```

#### Observation
- There is insufficient evidence to conclude that the
training program is effective at reducing swim times for Men’s 1500 Freestyle
(p = 0.01). The program, on average, decreased swim time by 7.26 seconds (95% CI
on difference: pre - post > 0: 2.15  Inf)."

### 2.3 Wilcoxon Signed-Rank Test

```{r}
wilcox.test(datap$diff, alternative = "greater",mu=0)
```

#### Observation
- For wilcox test, we have p-value=0.02 which is similar to t-test but less than the significant value 0.05. This suggest that, our initial hypothesis of improve in the swim times is incorrect. 

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
