---
title: "HW9 - Multiple Regression"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
library(ggplot2)
library(MASS)
library(Hmisc)
library(epiDisplay)
library(vcd)
library(mnormt)
library(MASS)
library(car)
library(ggpubr)
library(PairedData)
library(lmtest)
library(xtable)
library(faraway)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

### Problem 1 

```{r}
# Read the data
senic <- read.delim("SENIC.txt", header=FALSE,sep = "")
names(senic) <- c('id','stay','age','inf_risk', 'cul_r','xray_r','beds','med','region','daily_avg','nurses','services')
head(senic)

# lm for infection risk
lm1 <- lm(stay ~ inf_risk, data=senic)
summary(lm1)

# lm for available facilities
lm2 <- lm(stay ~ services, data=senic)
summary(lm2)

#lm for xray ratio
lm3 <- lm(stay ~ xray_r, data=senic)
summary(lm3)

# lm plots
par(mfrow = c(1,3))
plot(senic$stay ~ senic$inf_risk, cex = 1.5, cex.lab = 1.5, las = 1, cex.main = 1.5, xlab='Infection risk', ylab='average length of stay')
abline(lm1, lwd = 2, col = "red")
plot(senic$stay ~ senic$services, cex = 1.5, cex.lab = 1.5, las = 1, cex.main = 1.5, xlab='Services', ylab='average length of stay')
abline(lm2, lwd = 2, col = "red")
plot(senic$stay ~ senic$xray_r, cex = 1.5, cex.lab = 1.5, las = 1, cex.main = 1.5, xlab='X ray ratio', ylab='average length of stay')
abline(lm3, lwd = 2, col = "red")
```

#### Observation
- If we observe the p-values, we have values < 0.05 for three predictors. This says that the null hypothesis of having coefficients = 0 is rejected and linear relation between predictor and response holds good. This suggest the linearity with confidence (1- pvalue)%. 
- Plots of regression functions also shows the linear relationship provide a good fit for all the three predictors.


### Problem 2

```{r}
#Mean squared error
mse1 <- mean(lm1$residuals^2)
mse2 <- mean(lm2$residuals^2)
mse3 <- mean(lm3$residuals^2)
mse <- c(mse1,mse2,mse3)
r2 <- c(summary(lm1)$r.squared, summary(lm2)$r.squared, summary(lm3)$r.squared)
df <- data.frame(MSE=mse,R2=r2)
row.names(df) <- c('Infection risk','Facilities and Services','Chest X-ray ratio')
df
```

#### Observation
- R-square values are under 0.3 for all the three predictors. This shows that less than 30% of variance in dependent variable is explained by these univariate models.
- For the 'Infection risk' predictor, we have less mean square value and hightest Rsquare, so we can say out of three predictors, 'Infection risk' has the largest reduction in variablity of the average length of stay.

### Problem 3

```{r}
# Residual plots
par(mfrow = c(1, 3))
plot(fitted(lm1), residuals(lm1))
abline(h = 0)
plot(fitted(lm2), abs(residuals(lm2)))
abline(h = 0)
plot(fitted(lm3), abs(residuals(lm3)))
abline(h = 0)

par(mfrow = c(1, 3))
plot(lm1,which=c(1),main='Residuals vs Fitted lm1',caption='')
plot(lm2,which=c(1),main='Residuals vs Fitted lm1',caption='')
plot(lm3,which=c(1),main='Residuals vs Fitted lm1',caption='')

# Linearity of residuals
par(mfrow = c(1, 3))
qqnorm(residuals(lm1), main = "QQ Plot-Infection risk")
qqline(residuals(lm1))
shapiro.test(residuals(lm1))

qqnorm(residuals(lm2), main = "QQ Plot-Services")
qqline(residuals(lm2))
shapiro.test(residuals(lm2))

qqnorm(residuals(lm3), main = "QQ Plot-Xray ratio")
qqline(residuals(lm3))
shapiro.test(residuals(lm3))

# Homoscedasticity 
ncvTest(lm1)
ncvTest(lm2)
ncvTest(lm3)

# Correlation
dwtest(lm1)
dwtest(lm2)
dwtest(lm3)
```

#### Observation
- If we see the residual plots, model with infection risk is better compared to other two. 
- Normality: The p-values of shapiro test are less than the significant value < 0.05. Low p-values reject the null hypothesis of normality and that suggests that residuals are not following normal distribution. However, if we see the Qplots, we can see the normality, but presence of outliers could be the reason behind the poor p-values.
- Homoscedasticity: ncvtest has null hypothesis of constant variance for residuals. We have p-values less than the significant value 0.05. This shows the variance is changing basing on the change in the fitted valus. 
- Correlation: Dwtest has the null hypothesis  that the residuals from a linear regression are uncorrelated. Our test has p-values higher than 0.05 supports the null hypothesis suggesting that dependent and independent variables are not autocorrelated. DW=2 also suggests the no autocorrelation. 

### Problem 4

```{r}
# Halfnormal quantiles to see influencing points
par(mfrow = c(1, 3))
halfnorm(cooks.distance(lm1),main='Infection risk')
halfnorm(cooks.distance(lm2),main='Services')
halfnorm(cooks.distance(lm3),main='Chest Xray')

# Influencing points 
lm1I <- influence.measures(lm1)
which(apply(lm1I$is.inf, 1, any))

# Refit the model without the case 47 and 112
#senic1 <- senic[!(senic$id %in% c(47,112)),]
lm1_r <- lm(stay ~ inf_risk, data=senic, subset=-c(47,112))
str(summary(lm1_r))

# Prediction interval for the old model
senic[c('inf_risk','stay')][c(47,112),]
fitted.values(lm1)[c(47,112)]
newdata <- data.frame(inf_risk = c(6.5,5.9))
pio <-predict(lm1, newdata = newdata, interval = "prediction")
row.names(pio) <- c('47','112')
pio

# Prediction interval for the new model
pi <-predict(lm1_r, newdata = newdata, interval = "prediction")
row.names(pi) <- c('47','112')
pi
```

#### Observation
- The R2 value has improved from 2.8 to 3 for the refit model. The pr value of refit model is less than the original but not significantly less.
- Original 'Length of stay' values are Y=(19.56, 17.94) if their 'Risk of infection' values are X = (6.5,5.9). Using the refit model, we get the prediction intervals as (8.32, 13.3), (7.97, 12.9) for X respectively. 
- We can say that, original values do not fall under the bounds of the prediction intervals. 
- If we can compare the prediction intervals of original and refit model for 'Infection risk', we can say original model's interval are better than refit model. Removing the points (47,112) did not perform better. 

### Problem 5

```{r}
```

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
