---
title: "HW5 - Common Probability Distributions"
author: "Madhu Peduri"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    highlight: tango
    number_sections: yes
subparagraph: yes
header-includes: \usepackage{hyperref} \usepackage{framed,xcolor} \definecolor{shadecolor}{rgb}{0.99,
  0.95, 0.9} \renewcommand\arraystretch{1.5} \usepackage[none]{hyphenat} \raggedright
  \usepackage{amsmath, amsthm, amssymb, bm}
---

  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(graphics)
options(digits = 3)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```
## Discrete Distributions

### Bernoulli
The **Bernoulli distribution**, named for [Jacob Bernoulli](https://en.wikipedia.org/wiki/Jacob_Bernoulli), assigns probability to the outcomes of a single Bernoulli experiment---one where the only possible outcomes can be thought of as a "success" or a "failure" (e.g., a coin toss). Here, the random variable $x$ can take on the values 1 (success) with probability $p$, or 0 (failure) with probability $q = 1 - p$.  The plot below contains the pmf of two Bernoulli distributions.  The first (in gray) has a probability of success $p = 0.2$ and the second (in black) has a probability of success $p = 0.5$.  

```{r}
x <- 0:1
plot(x, dbinom(x, 1, 0.2), type = "h", ylab = "f(x)", ylim = c(0,1), lwd = 8, col = "dark gray",
      main = "Bernoulli(0.2)")
lines(x, dbinom(x, 1, 0.5), type = "h", lwd = 2, col = "black")
legend(0.7, 1.0, c("Bernoulli(0.2)","Bernoulli(0.5)"),col=c("dark gray","black"), lwd=c(8,2))

```

The Bernoulli experiment forms the foundation for many of the next discrete distributions.


### Binomial
The **binomial distribution** applies when we perform $n$ Bernoulli experiments and are interested in the total number of "successes" observed.  The outcome here, $y = \sum x_i$, where $P(x_i = 1) = p$ and $P(x_i = 0) = 1 - p$.  The plot below displays three binomial distributions, all for $n = 10$ Bernoulli trials: in gray, $p = 0.5$; in blue, $p = 0.1$; and in green, $p = 0.9$.  
```{r}
x <- seq(0, 10, 1)
plot(x, dbinom(x, 10, 0.5), type = "h", ylab = "f(x)", lwd = 8, col = "dark gray", ylim = c(0,0.5),
     main = "Binomial(10, 0.5) pmf")
lines(x, dbinom(x, 10, 0.1), type = "h", lwd = 2, col = "blue")
lines(x, dbinom(x, 10, 0.9), type = "h", lwd = 2, col = "green")
legend(3, 0.5, c("Binomial(10,0.1)","Binomial(10,0.5)","Binomial(10,0.9)"), col=c("blue","dark gray","green"),lwd=c(2,8,2))

```

We can see the shifting of probability from low values for $p = 0.1$ to high values for $p = 0.9$.  This makes sense, as it becomes more likely with $p = 0.9$ to observe a success for an individual trial.  Thus, in 10 trials, more successes (e.g., 8, 9, or 10) are likely.  For $p = 0.5$, the number of successes are likely to be around 5 (e.g., half of the 10 trials).

### Hypergeometric
The **Hypergeometric distribution** is a discrete distribution that describes the probability of $x$ successes in $n$ draws without replacement wherein each draw is either success or failure.  In contrast, the **binomial distribution** describes the probability of $x$ successes in $n$ draws with replacement. We can represent this using below notation

$p(x) = \dfrac{choose(m,x).choose(n,k-x)}{choose(m+n,k)}$ 

- In the example below, we have different set of parameters. 
- If we choose 2 balls from the urn - $x={(0,2),(1,1),(2,0)}$. 
- If we choose 5 balls from the urn with different number of colored balls - (7,3),$x={(2,3)(3,2),(4,1),(5,0)}$. 
- Probability distribution shifts as the number of success or failures changes.
- Probability distribution is symmetric. 

```{r}
x <- seq(0, 10, 1)
plot(x, dhyper(x, 5, 5, 2), type = "h", ylab = "f(x)", lwd = 8, col="blue",
    main = "Hypergeometric pmf")
lines(x, dhyper(x, 5, 5, 5), type = "h", ylab = "f(x)", lwd = 4, col="yellow")
lines(x, dhyper(x, 7, 3, 5), type = "h", ylab = "f(x)", lwd = 2, col="green")
lines(x, dhyper(x, 3, 7, 5), type = "h", ylab = "f(x)", lwd = 1, col="red")
par(cex = 0.5)
legend(7.5, 0.55, c("Hypergeometric (10,5,5,2)","Hypergeometric (10,5,5, 5)","Hypergeometric (10,7,3,5)","Hypergeometric (10,3,7,5)"), col=c("blue","yellow"," green","red"),lwd=c(8,4,2,1))
```

### Poisson
The **Poisson distribution** expresses the probability of a given number of events occurring in a fixed interval of time or space if these events occur with a known constant mean rate and independently of the time since the last event. The Poisson distribution can also be used for the number of events in other specified intervals such as distance, area or volume.

We use below function to derive the poisson distribution
$p(x) = \dfrac{\lambda^{x}e^{-\lambda}}{x!},\quad for \quad x=0,1,2 \dots n$

- For poission distribution, we use lambda as the parameter. It denotes the number of successes that we get with in the given time frame. 
- We can observe from the below distributions that, as the lamda increases, we get the maximum probability towards the right. That is, to have more number of success events, number of trials should be more. 

```{r}
x <- seq(0, 5, 1)
poisl <- dpois(x, 1)
plot(x, poisl, ylab = "f(x)", main = "Poisson pmf",pch=16,ylim=c(0,0.6),col="blue")
lines(x, poisl,col="blue")
text(x=which(poisl == max(poisl))[1],y=poisl[poisl == max(poisl)][1],"lambda=1",col="blue")

par(new=TRUE)
poisl <- dpois(x, 2)
plot(x, poisl, ylab = "f(x)", main = "Poisson pmf",pch=16,ylim=c(0,0.6),col="green")
lines(x, poisl,col="green")
text(x=which(poisl == max(poisl))[1],y=poisl[poisl == max(poisl)][1],"lambda=2",col="green")

par(new=TRUE)
poisl <- dpois(x, 3)
plot(x, poisl, ylab = "f(x)", main = "Poisson pmf",pch=16,ylim=c(0,0.6),col="red")
lines(x, poisl,col="red")
text(x=which(poisl == max(poisl))[1],y=poisl[poisl == max(poisl)][1],"lambda=3",col="red")
```

### Geometric
The **geometric distribution** gives the probability that the first occurrence of success requires k independent trials, each with success probability p. We use below function to derive the geometric distribution.

$Pr(X=k) = p(1-p)^{k-1}$

We can make below observations from below plots,

- As the initial success probability $p$ is high, the probability of getting a success event of certian failures decreases. 
- For example, after 3 falilures (x=3), the probablity of observing a success event decreases from 0.14 to 0.05 as the initial probability $p$ increases from 0.2 to 0.6.
- For a given success probability, as the number of failure trials in creases, the probability of observing a success decreases.

```{r}
x <- seq(0, 20, 1)
#plot(x, dgeom(x, 0.2), type = "h", ylab = "f(x)", lwd = 2,
#    main = "Geometric(0.2) pmf")
geom <- dgeom(x, 0.2)
plot(x, geom, ylab = "f(x)", main = "Geometric pmf",pch=16,ylim=c(0,0.6),col="blue")
lines(x, geom,col="blue")
text(x=which(geom == max(geom))[1],y=geom[geom == max(geom)][1],"p=0.2",col="blue")

par(new=TRUE)
geom <- dgeom(x, 0.4)
plot(x, geom, ylab = "f(x)", main = "Geometric pmf",pch=16,ylim=c(0,0.6),col="green")
lines(x, geom,col="green")
text(x=which(geom == max(geom))[1],y=geom[geom == max(geom)][1],"p=0.4",col="green")

par(new=TRUE)
geom <- dgeom(x, 0.6)
plot(x, geom, ylab = "f(x)", main = "Geometric pmf",pch=16,ylim=c(0,0.6),col="red")
lines(x, geom,col="red")
text(x=which(geom == max(geom))[1],y=geom[geom == max(geom)][1],"p=0.6",col="red")
```

### Negative Binomial
A **Negative binomial distribution** (also called the Pascal Distribution) is a discrete probability distribution, for random variables in a negative binomial experiment, that models the number of successes in a sequence of independent and identically distributed Bernoulli trials before a specified (non-random) number of failures (denoted r) occurs. We use below formula to derive this,

$f(x) = {(x+r-1)_{C}}_{r}(1-p)^{r}p^{x} \quad for \quad x=0,1,\dots$

We make below observations from below plots,

- We keep the initial probability of success, p=0.4, as constant and vary the number of successes $r$ to observe the pattern as number of trials are increased.
- As the number of successes $r$ increases, the highest probability decreases for a given number of trials. For example, for 20 trials as $r$ increases from 1 to 5, the highest probability $f(x)$ decreases from 0.4 to 0.08. 
- When $r=1$, it is a special case of geometric distribution. 
- As the number of successes $r$ increases, the number of trials should be more to get the highest probability. For example, As $r$ increases from 1 to 5, the number of trials needed increases from 0 to 5 to get the highest probability $f(x)$

```{r}
x <- seq(0, 20, 1)
nbinom <- dnbinom(x,1,0.4)
plot(x, nbinom, type = "p", ylab = "f(x)", pch=16,ylim=c(0,0.8),
     main = "Negative Binomial(0.4) pmf",col="red")
lines(x, nbinom,col="red")
text(x=which(nbinom == max(nbinom))[1],y=nbinom[nbinom == max(nbinom)][1]+0.04,"r=1",col="red")

par(new=TRUE)
nbinom <- dnbinom(x,3,0.4)
plot(x, nbinom, type = "p", ylab = "f(x)", pch=16,ylim=c(0,0.8),
     main = "Negative Binomial(0.4) pmf",col="blue")
lines(x, nbinom,col="blue")
text(x=which(nbinom == max(nbinom))[1],y=nbinom[nbinom == max(nbinom)][1]+0.04,"r=3",col="blue")

par(new=TRUE)
nbinom <- dnbinom(x,5,0.4)
plot(x, nbinom, type = "p", ylab = "f(x)", pch=16,ylim=c(0,0.8),
     main = "Negative Binomial(0.4) pmf",col="green")
lines(x, nbinom,col="green")
text(x=which(nbinom == max(nbinom))[1],y=nbinom[nbinom == max(nbinom)][1]+0.04,"r=5",col="green")
```

## Continuous Distributions

### Normal
The **Normal distribution** is a type of continuous probability distribution for a real-valued random variable. Below are some of the properties of this distribution,

- The mean, mode and median are all equal.
- The curve is symmetric at the center (i.e. around the mean).
 - Exactly half of the values are to the left of center and exactly half the values are to the right.
- The total area under the curve is 1.

We use below function to dervive this distribution,
\
$f(x) = \dfrac{1}{\sqrt{2 \pi \sigma^{2}}}e^{- \dfrac{(x-\mu)^{2}}{2\sigma^{2}}} \quad for \quad -\infty < x < \infty$

We can make below observations,

- As the standard deviation $\sigma$ increases, the curve flattens along the x-axis. suggesting, the values are scattered away from the mean.
- As the $\sigma$ increases, the probability $f(x)$ is decreased. 

```{r}
x <- seq(-5, 5, 0.01)
cp = c("red","blue","green","gray")
j <- 0
for (i in seq(1,4,0.8)){
  j <- j + 1
  nnorm <- dnorm(x, 0, i)
  plot(x, nnorm, type = "l",xlim=c(-5,5),ylim=c(0,0.6), ylab = "f(x)", main = "Normal pdf",col=cp[j])
  text(x=x[nnorm == max(nnorm)][1],y=nnorm[nnorm == max(nnorm)][1]+0.02,bquote(paste(sigma,"=",.(i),sep = "")),col=cp[j])
  par(new=TRUE)
}
```

### Exponential 
The **Exponential distribution** is a probability distribution that describes time between events in a Poisson process, a process in which events occur continuously and independently at a constant average rate. There is a strong relationship between the Poisson distribution and the Exponential distribution. For example, let’s say a Poisson distribution models the number of births in a given time period. The time in between each birth can be modeled with an exponential distribution. We use below function to represent this distribution,

$f(x) = \lambda e^{-\lambda x \quad for \quad x \in [0, \infty)}$

We can make below observations,

- In exponential distribution, $\lambda$ denotes the time between each event. 
- We can see that as the $\lambda$ increases from 0.5 to 2, the probability of a succuess event, $f(x)$, increases from 0.5 to 2. 
- We can also notice that, the probability of finding a success event reduces as we increase the number of trials. The rate of decrease in this probability is more as we increase the time between the events, that is $\lambda$

```{r}
x <- seq(0, 10, 0.01)
cp = c("red","blue","green","gray")
j <- 0
for (i in seq(0.5,2,0.5)){
  j <- j + 1
  nexp <- dexp(x, i)
  plot(x, nexp,type = "l",xlim=c(0,10),ylim=c(0,2), ylab = "f(x)", main = "Exponential pdf",col=cp[j])
  text(x=x[nexp == max(nexp)][1]+0.4,y=nexp[nexp == max(nexp)][1]-0.02,bquote(paste(lambda,"=",.(i),sep="")),col=cp[j])
  par(new=TRUE)
}
```

### Chisquare
The chi-square distribution with k degrees of freedom is the distribution of a sum of the squares of k independent standard normal random variables. We use below notation to represent this distribution,

If $Z_{1},\dots,Z_{k}$ are $k$ independent, standard normal random variables, chi-squared distribution is given by,
\
$Q = \sum\limits_{i=1}^{k}Z_{i}^{2}$

We can make below observations,

- $k$, degrees of freedom, is the only parameter in Chi-square distribution.
- As we increase the $k$, the distribution goes skew towards the right and follows normal distribution.  

```{r}
x <- seq(0, 20, 0.01)
j <- 0
cp = c("red","blue","green","gray","black","purple")
for (i in c(1,2,3,4,6,10)){
  j <- j + 1
  nchi <- dchisq(x, i)
  plot(x, nchi,type = "l",xlim=c(0,20),ylim=c(0,0.5), ylab = "f(x)", main = "Chi-square pdf",col=cp[j])
  par(new=TRUE)
}
legend( "topright", c("k=1", "k=2", "k=3", "k=4","k=6","k=10"), 
text.col=cp )
```

### Student's t

The **Student's t-distribution** is any member of a family of continuous probability distributions that arise when estimating the mean of a normally-distributed population in situations where the sample size is small and the population's standard deviation is unknown.The $t$ score is a ratio between the difference between two groups and the difference within the groups. The larger the t score, the more difference there is between groups. The smaller the t score, the more similarity there is between groups. We use below notation to represent this distribution,

$t(v) = \dfrac{\bar{x}-\mu}{\sqrt{\dfrac{s^{2}}{v}}}$

We can make below observations from the plots,

- As the degrees of freedom increases, we can see the Student's T distribution tend towards the normal distribution. 

```{r}
x <- seq(-5, 5, 0.01)
j <- 0
cp = c("red","blue","green","gray","black","purple","black")
for (i in c(1,2,3,4,6,10)){
  j <- j + 1
  nstud <- dt(x, i)
  plot(x, nstud,type = "l",xlim=c(-5,5),ylim=c(0,0.5), ylab = "f(x)", main = "Student's t pdf",col=cp[j])
  par(new=TRUE)
}
plot(x, dnorm(x, 0, 1), type = "l",lty=3,lwd=2,xlim=c(-5,5),ylim=c(0,0.5), ylab = "f(x)",col="black")
legend( "topright", c("k=1", "k=2", "k=3", "k=4","k=6","k=10","--- = normal"), 
text.col=cp )
```

### F

The F-distribution is a continuous probability distribution that arises frequently as the null distribution of a test statistic, most notably in the analysis of variance (ANOVA) and other F-tests. 

We can make below observations,

- If we increase either numerator or denominator, the plot tends towards the right with an increase in the probability. 

```{r}
x <- seq(0, 6, 0.01)
plot(x, df(x, 8, 15), type = "l", ylim=c(0,0.8), ylab = "f(x)", main = "F pdf",col="blue")

par(new=TRUE)
plot(x, df(x, 12, 15), type = "l",lty=3,ylim=c(0,0.8), ylab = "f(x)",col="blue")

par(new=TRUE)
plot(x, df(x, 10, 6), type = "l",ylim=c(0,0.8), ylab = "f(x)",col="red")

par(new=TRUE)
plot(x, df(x, 10, 8), type = "l",lty=3,ylim=c(0,0.8), ylab = "f(x)",col="red")

legend( "topright", c("blue=change in numerator","green=change in denominator","... = increase in value"), text.col=c("blue","red","black") )
```

### Beta

The **beta distribution** is a family of continuous probability distributions defined on the interval [0, 1] parameterized by two positive shape parameters, denoted by $\alpha$ and $\beta$, that appear as exponents of the random variable and control the shape of the distribution.

```{r}
x <- seq(0, 1, 0.01)
j <- 0
cp = c("red","blue","green","gray","purple")
alpha = c(0.5,5,1,2,2)
beta = c(0.5,1,3,2,5)
for (i in seq(1,5)){
  j <- j + 1
  nbeta <- dbeta(x, alpha[i],beta[i])
  plot(x, nbeta,type = "l",xlim=c(0,1),ylim=c(0,2.5), ylab = "f(x)", main = "Beta pdf",col=cp[j])
  par(new=TRUE)
}
legend( "topright", c("a=0.5,b=0.5","a=5,b=1","a=1,b=3","a=2,b=2","a=2,b=5"), 
text.col=cp )
```

### Logistic

The **logistic distribution** is a continuous probability distribution. Its cumulative distribution function is the logistic function, which appears in logistic regression and feedforward neural networks. It resembles the normal distribution in shape but has heavier tails (higher kurtosis).

```{r}
x <- seq(-5, 20, 0.01)
j <- 0
cp = c("red","blue","green","gray","purple")
loc = c(5,9,9,6,2)
scl = c(2,3,4,2,1)
for (i in seq(1,5)){
  j <- j + 1
  nlogis <- dlogis(x, loc[i],scl[i])
  plot(x, nlogis,type = "l",xlim=c(-5,20),ylim=c(0,0.3), ylab = "f(x)", main = "Logistic pdf",col=cp[j])
  par(new=TRUE)
}
legend( "topright", c("l=5,s=2","l=9,s=3","l=9,s=4","l=6,s=2","l=2,s=1"), 
text.col=cp )
```

## Document Information.

All of the statistical analyses in this document will be performed using `r R.version.string`.  R packages used will be maintained using the [packrat](http://rstudio.github.io/packrat/) dependency management system.  

```{r session-info}
sessionInfo()
```
